---
title: "EDA"
author: "Keith Williams"
date: "January 19, 2017"
output: 
    html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse) # data wrangling and visualization

# file should exist in the parent directory of the repo
# parent/
# |-> 10kDiabetes.csv
# |-> kdd-diabetes/
# |   |-> R/
file_name <- "../../10kDiabetes.csv"

# read file
diabetes <- read_csv(file_name, na = c("?", ""))
```

## Introduce the rows and columns  

```{r}
# view structure
glimpse(diabetes)

# outcome variable
mean(diabetes$readmitted)
```

There are 10,000 records, 39.65% of which were readmitted within 30 days. Features include demographics, admission conditions, procedure and lab descriptions and results, medication changes, and diagnoses.  

## Use descriptive statistics to describe the data  

```{r, fig.height = 6}
# proportion of missing records for each variable
prop_na <- map_dbl(diabetes, ~mean(is.na(.x)))
data_frame(variable = names(prop_na), proportion_na = prop_na) %>% 
    ggplot(aes(forcats::fct_reorder(variable, prop_na), prop_na)) +
    geom_bar(stat = "identity") +
    labs(x = "variable", y = "proportion missing values", title = "Missing Values") +
    coord_flip()
```

Weight, payer code, and medical specialty have so many missing values, that they likely won't be useful. It may be possible to impute admission_source_id, admission_type_id, discharge_disposition_id, race, and the diagnoses.  

```{r, fig.height = 8}
# demographic distributions
race_plot <- ggplot(diabetes, aes(race, fill = readmitted)) + geom_bar()
age_plot <- ggplot(diabetes, aes(age, fill = readmitted)) + geom_bar()
gender_plot <- ggplot(diabetes, aes(gender, fill = readmitted)) + geom_bar()

gridExtra::grid.arrange(race_plot, age_plot, gender_plot)
```

```{r, fig.height = 7, fig.width = 9}
# summarise numeric columns
numeric_cols <- diabetes[, map_lgl(diabetes, ~is.numeric(.x))] %>% 
    select(-rowID)
summary(numeric_cols)

# look at numeric column distributions by readmission status
tidy_numeric_cols <- numeric_cols %>% 
    bind_cols(select(diabetes, readmitted)) %>% 
    gather("variable", "value", -readmitted)

# boxplots
ggplot(tidy_numeric_cols, aes(factor(0), value)) + 
    geom_boxplot() + 
    facet_wrap(~variable, scales = "free_x") +
    labs(x = NULL, title = "Numeric Column Boxplots") +
    coord_flip() +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

# Distribution by readmission status
ggplot(tidy_numeric_cols, aes(value, color = readmitted, fill = readmitted)) +
    geom_density(alpha = 0.4) + 
    facet_wrap(~variable, scales = "free")
```

```{r, fig.width = 9, fig.height = 11}
# medications
select(diabetes, max_glu_serum:readmitted) %>% 
    gather("med", "description", -readmitted) %>% 
    ggplot(aes(description, fill = readmitted)) +
    geom_bar() +
    facet_wrap(~med, scales = "free_x", ncol = 4)
```

One can see that some the medications have zero or near-zero variance:  
    - acarbose  
    - acetohexamide  
    - chlorpropamide
    - citoglipton  
    - examide  
    - glimepiride:pioglitazone  
    - glipizide.metformin  
    - metformin.pioglitazone  
    - metformin.rosiglitazone  
    - miglitol  
    - tolazamide  
    - tolbutamide  
    - troglitazone  
    
These medications will not be useful in predicting readmission with little to no variance among records.    

Of particular interest, it appears "change" has a much higher proportion of readmissions than no change:  
```{r}
ggplot(diabetes, aes(change, fill = readmitted)) + 
    geom_bar(position = "fill")
```

Also, A1Cresult seems to have an association with readmission:  
```{r}
diabetes %>% 
    mutate(A1C_tested = ifelse(A1Cresult == "None", "not_tested", "tested")) %>% 
ggplot(aes(A1C_tested, fill = readmitted)) +
    geom_bar(position = "fill")
```

```{r, fig.width = 9, fig.height = 6}
# look for commonalities in primary diagnosis
stringr::str_replace_all(diabetes$diag_1_desc, pattern = "unspecified", "") %>% 
wordcloud::wordcloud(max.words = 250, 
                     random.color = FALSE, 
                     colors = viridis::viridis(n = 7))
```

