---
title: "EDA"
author: "Keith Williams"
date: "January 19, 2017"
output: 
    html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)
```

```{r load}
library(tidyverse) # data wrangling and visualization

# file should exist in the parent directory of the repo
# parent/
# |-> 10kDiabetes.csv
# |-> kdd-diabetes/
# |   |-> R/
file_name <- "../../10kDiabetes.csv"

# read file
diabetes <- read_csv(file_name, na = c("?", ""))
```

## Introduce the rows and columns  

```{r structure}
# view structure
glimpse(diabetes)

# outcome variable
mean(diabetes$readmitted)
```

There are 10,000 records, 39.65% of which were readmitted within 30 days. Features include demographics, admission conditions, procedure and lab descriptions and results, medication changes, and diagnoses.  

## Use descriptive statistics to describe the data  

```{r missing values, fig.height = 6}
# proportion of missing records for each variable
prop_na <- map_dbl(diabetes, ~mean(is.na(.x)))
data_frame(variable = names(prop_na), proportion_na = prop_na) %>% 
    ggplot(aes(forcats::fct_reorder(variable, prop_na), prop_na)) +
    geom_bar(stat = "identity") +
    labs(x = "variable", y = "proportion missing values", title = "Missing Values") +
    coord_flip()
```

Weight, payer code, and medical specialty have so many missing values, that they likely won't be useful. It may be possible to impute admission_source_id, admission_type_id, discharge_disposition_id, race, and the diagnoses.  

```{r demographics, fig.height = 8}
# demographic distributions
race_plot <- ggplot(diabetes, aes(race, fill = readmitted)) + geom_bar()
age_plot <- ggplot(diabetes, aes(age, fill = readmitted)) + geom_bar()
gender_plot <- ggplot(diabetes, aes(gender, fill = readmitted)) + geom_bar()

gridExtra::grid.arrange(race_plot, age_plot, gender_plot)
```

```{r numeric distributions, fig.height = 7, fig.width = 9}
# summarise numeric columns
numeric_cols <- diabetes[, map_lgl(diabetes, ~is.numeric(.x))] %>% 
    select(-rowID)
summary(numeric_cols)

# look at numeric column distributions by readmission status
tidy_numeric_cols <- numeric_cols %>% 
    bind_cols(select(diabetes, readmitted)) %>% 
    gather("variable", "value", -readmitted)

# boxplots
ggplot(tidy_numeric_cols, aes(factor(0), value)) + 
    geom_boxplot() + 
    facet_wrap(~variable, scales = "free_x") +
    labs(x = NULL, title = "Numeric Column Boxplots") +
    coord_flip() +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

# Distribution by readmission status
ggplot(tidy_numeric_cols, aes(value, color = readmitted, fill = readmitted)) +
    geom_density(alpha = 0.4) + 
    facet_wrap(~variable, scales = "free")
```

```{r medications, fig.width = 9, fig.height = 11}
# medications
select(diabetes, max_glu_serum:readmitted) %>% 
    gather("med", "description", -readmitted) %>% 
    ggplot(aes(description, fill = readmitted)) +
    geom_bar() +
    facet_wrap(~med, scales = "free_x", ncol = 4)
```

One can see that some the medications have zero or near-zero variance:  
    - acarbose  
    - acetohexamide  
    - chlorpropamide
    - citoglipton  
    - examide  
    - glimepiride:pioglitazone  
    - glipizide.metformin  
    - metformin.pioglitazone  
    - metformin.rosiglitazone  
    - miglitol  
    - tolazamide  
    - tolbutamide  
    - troglitazone  
    
These medications will not be useful in predicting readmission with little to no variance among records.    

Of particular interest, it appears "change" has a much higher proportion of readmissions than no change:  
```{r change meds}
ggplot(diabetes, aes(change, fill = readmitted)) + 
    geom_bar(position = "fill")
```

Also, A1Cresult seems to have an association with readmission:  
```{r A1Cresult}
diabetes %>% 
    mutate(A1C_tested = ifelse(A1Cresult == "None", "not_tested", "tested")) %>% 
ggplot(aes(A1C_tested, fill = readmitted)) +
    geom_bar(position = "fill")
```

```{r diagnosis wordcloud, fig.width = 9, fig.height = 6}
# look for commonalities in primary diagnosis
stringr::str_replace_all(diabetes$diag_1_desc, pattern = "unspecified", "") %>% 
wordcloud::wordcloud(max.words = 250, 
                     random.color = FALSE, 
                     colors = viridis::viridis(n = 7))
```

## Prepare Modeling Dataset

```{r}
# dropping columns due to missingness, near zero variance, and irrelevance for modeling (IDs)
for_model <- diabetes %>% 
    select(-c(rowID,
              weight,
              payer_code,
              medical_specialty,
              diag_1,
              diag_1_desc,
              diag_2,
              diag_2_desc,
              diag_3,
              diag_3_desc,
              acarbose,
              acetohexamide,
              chlorpropamide,
              citoglipton,
              examide,
              glimepiride.pioglitazone,
              glipizide.metformin,
              metformin.pioglitazone,
              metformin.rosiglitazone,
              miglitol,
              nateglinide,
              tolazamide,
              tolbutamide,
              troglitazone
              )) %>% 
    # fill missing admission source and types and discharge disposition with 'Not Available'
    tidyr::replace_na(list(admission_source_id = "Not Available", 
                           admission_type_id = "Not Available",
                           discharge_disposition_id = "Not Available")) %>% 
    # map A1C results into tested and not
    mutate(A1C_tested = ifelse(A1Cresult == "None", "not_tested", "tested")) %>% 
    select(-A1Cresult)
    
```

```{r}
for_model %>% 
    as.data.frame() %>% 
    saveRDS('../../data_for_modeling.rds')
```

```{r impute race}
# convert charcters to factors to prepare for KNN  
for_knn <- for_model %>% 
    mutate_if(is.character, as.factor)

# split into data to train KNN on race, remove response variable
train <- for_knn %>% 
    filter(!is.na(race)) %>% 
    select(-readmitted)

test <- for_knn %>% 
    filter(is.na(race)) %>% 
    select(-race, -readmitted)

# fit KNN model, k = 3
knn_mdl <- caret::knn3(race ~ ., data = train, k = 3)
# make predictions for missing race observations
race_imputations <- caret::predict.knn3(knn_mdl, newdata = test, type = "class")

# recombine the train and newly imputed test sets
test_imputed <- for_knn %>% 
    filter(is.na(race))
test_imputed$race <- race_imputations
full_imputed <- bind_rows(filter(for_knn, !is.na(race)), test_imputed)
```

```{r dummy categorical}
# library(caret)
# diabetes_dummy <- dummyVars(~ race + gender + age + admission_type_id + max_glu_serum +
#                             A1Cresult , data = diabetes, levelsOnly = TRUE,fullRank = T)
# frame <- data.frame( predict(diabetes_dummy, newdata = diabetes ) )
# for_model <- cbind(for_model,frame)
# glimpse(for_model)

dummied <- model.matrix(readmitted ~ ., data = full_imputed)
```

```{r correlation heatmap}
numeric_cols <- names(full_imputed)[map_lgl(full_imputed, is.numeric)]

cormat <- full_imputed %>%
    select(one_of(numeric_cols), readmitted) %>% 
    cor(use = "complete.obs")
cormat[upper.tri(cormat)] <- NA

cormat %>% 
    reshape2::melt(na.rm = TRUE) %>%
    ggplot(aes(Var2, Var1, fill = value)) +
    geom_tile(color = 'gray90') +
    scale_fill_gradient2(low = "#67a9cf", high = "#ef8a62", limit = c(-1, 1), name = "correlation") +
    coord_equal() +
    theme(axis.text.x = element_text(size = 14,angle = 45, vjust = 1, hjust = 1),
          axis.text.y = element_text(size = 14),
          axis.title = element_blank(),
          panel.background = element_blank()) +
    ggtitle("Correlation Matrix")
```

```{r}
# data_for_model <- readRDS('../../data_for_modeling.rds')
```

```{r pearsons correlation analysis}
proceduresData <- subset(for_model, select =  c(numeric_cols, "readmitted"))
matrix <- as.matrix(proceduresData)
Hmisc::rcorr(matrix,type="pearson")
```

```{r chi-square analysis, eval = F}
library(gmodels)

CrossTable(for_model$glimepiride, for_model$readmitted, chisq = TRUE, expected = TRUE, format = "SPSS")
```


```{r chi-square test, eval = F}

CrossTable( for_model$readmitted , for_model$insulin , chisq = TRUE, expected = TRUE, format = "SPSS")
CrossTable( for_model$readmitted , for_model$A1Cresult , chisq = TRUE, expected = TRUE, format = "SPSS")

```

```{r PCA}
## PCA
pca <- prcomp(dummied)
s <- summary(pca)
pca_summary <- data_frame(pc = 1:91, prop_var = s$importance[2, ]) %>%
    mutate(cumvar = cumsum(prop_var))

pve <- ggplot(pca_summary, aes(pc, prop_var)) +
    geom_line() +
    geom_point() +
    # scale_x_continuous(breaks = 0:91) +
    theme(panel.grid.minor.x = element_blank()) +
    labs(x = "Principal Component",
         y = "Proportion of Variance Explained")

cve <- ggplot(pca_summary, aes(pc, cumvar)) + 
    geom_line() + 
    # scale_x_continuous(breaks = 0:91) +
    theme(panel.grid.minor.x = element_blank()) +
    labs(x = "Principal Component",
         y = "Cumulative Variance Explained",
         title = "Principal Components Analysis")

gridExtra::grid.arrange(cve, pve)
```

```{r}
data_frame(PC1 = pca$x[,1], PC2 = pca$x[,2], readmitted = full_imputed$readmitted) %>% 
    ggplot(aes(PC1, PC2, color = readmitted)) + 
    geom_point(alpha = 0.3) +
    scale_color_manual(values = c("#545454", "#E34949")) +
    guides(color = guide_legend(override.aes = list(alpha = 1))) +
    ggtitle("First Two Principal Components")
```

```{r}
data_frame(PC1 = pca$x[,1], PC2 = pca$x[,2], readmitted = full_imputed$readmitted) %>% 
    gather("PC", "value", PC1, PC2) %>% 
    ggplot(aes(value, fill = readmitted, color = readmitted)) +
    geom_density(alpha = 0.3) +
    facet_wrap(~PC, ncol = 1, scales = "free_y") +
    scale_color_manual(values = c("#545454", "#E34949")) +
    scale_fill_manual(values = c("#545454", "#E34949")) +
    ggtitle("Distribution of First Principal Component")
```

```{r logistic regression}
mdl <- glm(readmitted ~ ., data = full_imputed)
summary(mdl)
broom::tidy(mdl) %>% 
    filter(p.value < 0.2) %>% 
    mutate(odds_ratio = exp(estimate),
           readmission = ifelse(odds_ratio > 1, "More Likely", "Less Likely")) %>% 
    ggplot(aes(reorder(term, odds_ratio), 
               odds_ratio, 
               fill = readmission, 
               alpha = -p.value, 
               color = readmission)) +
    geom_bar(stat = "identity") +
    scale_color_manual(values = c("#545454", "#E34949")) +
    scale_fill_manual(values = c("#545454", "#E34949")) +
    labs(x = "Odds Ratio",
         y = "Feature",
         alpha = "significance",
         title = "Most Important Features",
         subtitle = "Compared to race:AA, age:[0-10], admitted:Elective") +
    guides(alpha = guide_legend(label = F)) +
    coord_flip()
```

## Prediction Task  

Going to use full, imputed data, split into train and test, then train a few algorithms to predict.  

```{r}
set.seed(666)
# split into train and test using stratified split
full_imputed_factor <- full_imputed %>% 
    mutate(readmitted = as.factor(readmitted))
inTrain <- caret::createDataPartition(full_imputed_factor$readmitted, p = 0.8, list = FALSE)
trainXy <- full_imputed_factor[inTrain,]
testXy <- full_imputed_factor[-inTrain,]
```

```{r}
gbm_grid <- expand.grid(interaction.depth = c(1, 2, 3, 4, 5),
                        n.trees = (2:8)*25,
                        shrinkage = c(0.03, 0.1),
                        n.minobsinnode = 20)
fit_control <- caret::trainControl(method = "repeatedcv", number = 5, repeats = 5)
gbm_fit <- caret::train(readmitted ~ ., data = trainXy,
                        method = 'gbm',
                        trControl = fit_control,
                        tuneGrid = gbm_grid,
                        verbose = F)
```

```{r}
ggplot(gbm_fit) +
    ggtitle("Cross Validation Accuracy Over Tuning Parameters")
```

```{r}
# testing accuracy
yhat <- predict(gbm_fit, newdata = testXy)
caret::confusionMatrix(yhat, testXy$readmitted)
```

```{r}
svm_fit <- caret::train(readmitted ~ ., data = trainXy,
                        method = "svmRadial",
                        trControl = fit_control,
                        preProc = c("center", "scale"),
                        tuneLength = 4,
                        verbose = F)
```

```{r}
ggplot(svm_fit) +
    ggtitle("Cross Validation Accuracy Over Tuning Parameters")
```

```{r}
# testing accuracy
yhat <- predict(svm_fit, newdata = testXy)
caret::confusionMatrix(yhat, testXy$readmitted)
```

